---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "10/29/2020"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

```

```{r setup}
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(parallel)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(patchwork)
library(abind)
library(INLA)
library(gsubfn)
library(dplyr)
library(RCurl)
library(dplyr)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(rjags)
library(INLA)
library(lme4)
library(pbapply)
#library(jsonlite)
set.seed(123)

source('./R/read_dependent_scripts.R')

std.pop.age.region.sex.race <- readRDS('./Data/std.pop.age.region.sex.race.rds')
std.pop.age.sex.race <- readRDS('./Data/std.pop.age.sex.race.rds')

```

Read in VA data

```{r}
v1 <- read_csv('./Data/Confidential/aggdata_withregion.csv')
v1$race_recode[is.na(v1$race_recode)] <- 999
v1$pop <- v1$n #pop is specific to the agec, region, sex, year but does NOT vary by race or quarter

v1$N_deaths[v1$N_deaths=='<10'] <- 5 #just as a placeholder...can either have liam run the analysis locally to avoid masking, or figure out an imputation method
v1$N_deaths <- as.numeric(v1$N_deaths)

  v1$month <- NA
  v1$month[v1$qtr==1] <- 1
  v1$month[v1$qtr==2] <- 4
  v1$month[v1$qtr==3] <- 7
  v1$month[v1$qtr==4] <- 10
  v1$date <- as.Date(paste(v1$year, v1$month, '01', sep='-'))

  v1 <- v1[v1$date<=as.Date('2021-01-01'),]
  
  v1 <- v1[!is.na(v1$region) & !is.na(v1$sex) & !is.na(v1$agec),] # for now, ignore deaths where sex, or region or age is NA
  
  summary.va <- v1[v1$year<2020,] %>%
    group_by(agec, race_recode, region) %>%
    summarize(N_deaths=sum(N_deaths),pop=mean(n)) %>%
    mutate(inc_va=N_deaths/pop*100000) %>%
    select(agec, race_recode, region, inc_va)

  v1$source <- 'va'
  
  v1$time <- round(as.numeric(v1$date - min(v1$date))/30.3)
  v1$time_scale <- as.vector(scale(v1$time)) 

  
   v1$qtr2 <- 0
  v1$qtr2[v1$qtr==2] <- 1
  v1$qtr3 <- 0
  v1$qtr3[v1$qtr==3] <- 1
  v1$qtr4 <- 0
  v1$qtr4[v1$qtr==4] <- 1
  
    v1$subgroup_combo <- as.factor(paste(v1$race_recode, v1$agec, v1$sex,v1$region))

      v1$all_cause_pre <- v1$N_deaths
  v1$all_cause_pre[v1$date >= '2020-01-01'] <- NA #bc it is quarterly data, Q1 includes pandeic period
  v1$race_recode <- as.factor(v1$race_recode)
  
  v1 <- v1[v1$agec !="Under 25 Years",]

    v1$inc <- v1$N_deaths/v1$pop*100000
    v1$log_pop <- log(v1$pop/100000)
    
    v1 <- v1[v1$year <=2020,]
```

## Compare VA group incidence with cdc
```{r}
summary.cdc <- readRDS('./Data/grp_inc_cdc.R')
comp_inc <- merge(summary.cdc,summary.va, by=c('agec', 'race_recode' ,'region'))

comp_inc

write.csv(comp_inc, './Data/compare_baseline_inc.csv')

p0 <- ggplot(comp_inc,aes(x=inc_cdc, y=inc_va)) +
  geom_point() +
  theme_classic() +
      facet_wrap(~race_recode,  nrow=1) +
  geom_abline(slope=1, intercept=0)

p0
```


```{r, fig.width=8, fig.height=8}
plot.fun(ds= v1[v1$sex=='M' &v1$race_recode=='1',], xvar='date', yvar='inc')

plot.fun(ds= v1[v1$sex=='F' &v1$race_recode=='1',], xvar='date', yvar='inc')

plot.fun(ds= v1[v1$sex=='M' &v1$race_recode=='2',], xvar='date', yvar='inc')
```

https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-detail.html
Race:
1 = White Alone
2 = Black or African American Alone
3 = American Indian or Alaska Native Alone
4 = Asian Alone
5 = Native Hawaiian and Other Pacific Islander Alone
6 = Two or more races

Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic


## Model


```{r}

mod1 <- pblapply(combine.data, inla_model2) #Maybe update this
res1 <- bind_rows( mod1$va$preds_covars.m) #add VA data too
res1$year <- year(res1$date)
```


We can then take the results from the INLA and aggregate over different groups. For example, by age, by age/race, by age/region, by age/race/region:
```{r}
preds.age <- res1 %>%
  group_by(source,agec, date, variable) %>%
  summarize_grps_sums %>%
  ungroup() %>%
  group_by(source,agec, date) %>%
  summarize_grps_quantiles

preds.age_race <- res1 %>%
  group_by(source,agec, race_recode, date, variable) %>%
  summarize_grps_sums %>%
  ungroup() %>%
  group_by(source,agec,race_recode, date) %>%
  summarize_grps_quantiles

preds.age.region <- res1 %>%
  group_by(source,agec,region, date, variable) %>%
  summarize_grps_sums %>%
  ungroup() %>%
  group_by(source,agec,region, date) %>%
  summarize_grps_quantiles

preds.age.race.region <- res1 %>%
  group_by(source,agec,region,race_recode, date, variable) %>%
  summarize_grps_sums %>%
  ungroup() %>%
  group_by(source,agec,region,race_recode, date) %>%
  summarize_grps_quantiles

#Across multiple time points--this needs
preds.age.yr <- res1 %>%
  group_by(source,agec, year, variable) %>%
  summarize_grps_sums %>%
  ungroup() %>%
  group_by(source,agec, year) %>%
  summarize_grps_quantiles

preds.age.yr <- res1 %>%
  group_by(source, year, variable) %>%
  summarize_grps_sums %>%
  ungroup() %>%
  group_by(source, year) %>%
  summarize_grps_quantiles
```

```{r}
#For creating standardized summaries of RR and incidence, start at most disaggregate level
preds.age.race.region.sex.qtr <- res1[res1$year==2020,] %>%
  group_by(source,agec,region,race_recode,sex, date, variable) %>%
  summarize_grps_sums %>%
  ungroup() %>%
  group_by(source,agec,region,race_recode,sex, date) %>%
  summarize_grps_quantiles
```

```{r, fig.width=8, fig.height=6}
p1 <- obs_exp_plot(ds=preds.age[preds.age$source=='va',], xvar='date',yvar='N_deaths', groupvar='agec', colvar='agec')
p1 + facet_wrap(~agec, scales='free')

```

```{r, fig.width=8, fig.height=6}
p1 <- RR_plot(ds=preds.age[preds.age$source=='va',],  groupvar='agec', colvar='agec')
p1 + facet_wrap(~agec, scales='free')

```

or by age and race

```{r, fig.width=8, fig.height=6}

p1 <- obs_exp_plot(ds=preds.age_race[preds.age_race$source=='va',], xvar='date',yvar='N_deaths', groupvar='agec', colvar='agec')

p1 + facet_grid(agec~race_recode, scales='free')

```


```{r, fig.width=8, fig.height=6}

p1 <- RR_plot(ds=preds.age_race[preds.age_race$source=='va',],  groupvar='agec', colvar='agec', datemin='2018-01-01') 


p1 + facet_grid(agec~race_recode, scales='free')

```
Age and region 
```{r, fig.width=8, fig.height=6}
p1 <- obs_exp_plot(ds=preds.age.region[preds.age.region$source=='va',], xvar='date',yvar='N_deaths', groupvar='agec', colvar='agec')
p1 + facet_grid(agec~region, scales='free')

```

```{r, fig.width=8, fig.height=6}
p1 <- RR_plot(ds=preds.age.region[preds.age.region$source=='va',],  groupvar='agec', colvar='agec', datemin='2018-01-01') 
p1 + facet_grid(agec~region, scales='free')

```