---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "10/29/2020"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
params:
  agg.level: 'state'
  n.days.filter: 20
  web.version: FALSE
  extrap.date: '2020-01-26'
  count.start.date: '2020-03-01'
  end.data.date: '2020-09-05'
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

```

```{r setup}
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(parallel)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(patchwork)
library(abind)
library(INLA)
library(gsubfn)
library(dplyr)
library(RCurl)
library(dplyr)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(INLA)
library(lme4)
library(pbapply)
#library(jsonlite)
set.seed(123)

source('./R/read_dependent_scripts.R')

std.pop.age.region.sex.race <- readRDS('./Data/std.pop.age.region.sex.race.rds')
std.pop.age.sex.race <- readRDS('./Data/std.pop.age.sex.race.rds')
```

Popsize for the CDC data
```{r}
filled_pop2 <- census_popsize_format()

```

Setup CDC data and merge with popsize
```{r}
ag3 <- cdc_data_setup_multilevel()
ag3$log_pop <- log(ag3$pop.interpol/100000)

ag3 <- ag3[ag3$agec !="Under 25 Years" ,]
  # summary.cdc <- ag3[ag3$year<2020,] %>%
  #   group_by(agec, race_recode, region) %>%
  #   summarize(N_deaths=sum(N_deaths),pop=mean(pop.interpol, na.rm=T)) %>%
  #   mutate(inc_cdc=N_deaths/pop*100000) %>%
  #   select(agec, race_recode, region, inc_cdc)
  # 
  # saveRDS(summary.cdc,'./Data/grp_inc_cdc.R')

```

https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-detail.html
Race:
1 = White Alone
2 = Black or African American Alone
3 = American Indian or Alaska Native Alone
4 = Asian Alone
5 = Native Hawaiian and Other Pacific Islander Alone
6 = Two or more races

Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic

 01 ... Under 1 year (includes not stated infant ages)
 02 ... 1 - 4 years
 03 ... 5 - 14 years
 04 ... 15 - 24 years
 05 ... 25 - 34 years
 06 ... 35 - 44 years
 07 ... 45 - 54 years
 08 ... 55 - 64 years
 09 ... 65 - 74 years
 10 ... 75 - 84 years
 11 ... 85 years and over
 12 ... Age not stated



## Fit Model


Fit to VA and CDC data separately. This function saves to an rds file summary tables with the excess incidence and RR by group (excluding raw counts and denominators). It also outputs the full summary tables that can be used in plotting in the list out.ds
```{r}
out.ds <- call_inla(label='cdc', ds=ag3)

```

```{r}
fit.plots <- mod.fit.plots()
rr.plot1 <- rr.plots()
saveRDS(fit.plots, './outputs/fit.plots.cdc.rds')
saveRDS(rr.plot1, './outputs/rr.plots.cdc.rds')

```


```{r, fig.width=8, fig.height=6}
fit.plots$p1
fit.plots$p3
fit.plots$p5

```

```{r, fig.width=8, fig.height=6}
rr.plot1$p2
rr.plot1$p4
rr.plot1$p6

```

```{r}

```

