---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "10/29/2020"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

```

```{r setup}
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(parallel)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(patchwork)
library(abind)
library(INLA)
library(gsubfn)
library(dplyr)
library(RCurl)
library(dplyr)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(rjags)
library(INLA)
library(lme4)
library(pbapply)
#library(jsonlite)
set.seed(123)

source('./R/read_dependent_scripts.R')

std.pop.age.region.sex.race <- readRDS('./Data/std.pop.age.region.sex.race.rds')
std.pop.age.sex.race <- readRDS('./Data/std.pop.age.sex.race.rds')

```

Read in VA data

```{r}
v1 <- read_csv('./Data/Confidential/aggdata_withregion.csv')
v1$race_recode[is.na(v1$race_recode)] <- 999
v1$pop <- v1$n #pop is specific to the agec, region, sex, year but does NOT vary by race or quarter

v1$N_deaths[v1$N_deaths=='<10'] <- 5 #just as a placeholder...can either have liam run the analysis locally to avoid masking, or figure out an imputation method
v1$N_deaths <- as.numeric(v1$N_deaths)

  v1$month <- NA
  v1$month[v1$qtr==1] <- 1
  v1$month[v1$qtr==2] <- 4
  v1$month[v1$qtr==3] <- 7
  v1$month[v1$qtr==4] <- 10
  v1$date <- as.Date(paste(v1$year, v1$month, '01', sep='-'))

  v1 <- v1[v1$date<=as.Date('2021-01-01'),]
  
  v1 <- v1[!is.na(v1$region) & !is.na(v1$sex) & !is.na(v1$agec),] # for now, ignore deaths where sex, or region or age is NA
  
  summary.va <- v1[v1$year<2020,] %>%
    group_by(agec, race_recode, region) %>%
    summarize(N_deaths=sum(N_deaths),pop=mean(n)) %>%
    mutate(inc_va=N_deaths/pop*100000) %>%
    select(agec, race_recode, region, inc_va)

  v1$source <- 'va'
  
  v1$time <- round(as.numeric(v1$date - min(v1$date))/30.3)
  v1$time_scale <- as.vector(scale(v1$time)) 

  
   v1$qtr2 <- 0
  v1$qtr2[v1$qtr==2] <- 1
  v1$qtr3 <- 0
  v1$qtr3[v1$qtr==3] <- 1
  v1$qtr4 <- 0
  v1$qtr4[v1$qtr==4] <- 1
  
    v1$subgroup_combo <- as.factor(paste(v1$race_recode, v1$agec, v1$sex,v1$region))

      v1$all_cause_pre <- v1$N_deaths
  v1$all_cause_pre[v1$date >= '2020-01-01'] <- NA #bc it is quarterly data, Q1 includes pandeic period
  v1$race_recode <- as.factor(v1$race_recode)
  
  v1 <- v1[v1$agec !="Under 25 Years",]

    v1$inc <- v1$N_deaths/v1$pop*100000
    v1$log_pop <- log(v1$pop/100000)
    
    v1 <- v1[v1$year <=2020,]
```

## Compare VA group incidence with cdc
```{r}
summary.cdc <- readRDS('./Data/grp_inc_cdc.R')
comp_inc <- merge(summary.cdc,summary.va, by=c('agec', 'race_recode' ,'region'))

comp_inc

write.csv(comp_inc, './Data/compare_baseline_inc.csv')

p0 <- ggplot(comp_inc,aes(x=inc_cdc, y=inc_va)) +
  geom_point() +
  theme_classic() +
      facet_wrap(~race_recode,  nrow=1) +
  geom_abline(slope=1, intercept=0)

p0
```


```{r, fig.width=8, fig.height=8}
plot.fun(ds= v1[v1$sex=='M' &v1$race_recode=='1',], xvar='date', yvar='inc')

plot.fun(ds= v1[v1$sex=='F' &v1$race_recode=='1',], xvar='date', yvar='inc')

plot.fun(ds= v1[v1$sex=='M' &v1$race_recode=='2',], xvar='date', yvar='inc')
```

https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-detail.html
Race:
1 = White Alone
2 = Black or African American Alone
3 = American Indian or Alaska Native Alone
4 = Asian Alone
5 = Native Hawaiian and Other Pacific Islander Alone
6 = Two or more races

Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic



## Fit Model


Fit to VA and CDC data separately. This function saves to an rds file summary tables with the excess incidence and RR by group (excluding raw counts and denominators). It also outputs the full summary tables that can be used in plotting in the list out.ds
```{r}
out.ds <- call_inla(label='va', ds=v1)

```

```{r}
fit.plots <- mod.fit.plots()
rr.plot1 <- rr.plots()
```


```{r, fig.width=8, fig.height=6}
fit.plots$p1
fit.plots$p3
fit.plots$p5

```

```{r, fig.width=8, fig.height=6}
rr.plot1$p2
rr.plot1$p4
rr.plot1$p6

```

```{r}

```

