---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "2/5/2022"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
params:
  agg.level: 'state'
  n.days.filter: 20
  web.version: FALSE
  extrap.date: '2020-01-26'
  count.start.date: '2020-03-01'
  end.data.date: '2020-09-05'
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

```

```{r setup}
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(parallel)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(patchwork)
library(abind)
library(INLA)
library(gsubfn)
library(dplyr)
library(RCurl)
library(dplyr)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(INLA)
library(lme4)
library(pbapply)
#library(jsonlite)
set.seed(123)

source('./R/read_dependent_scripts.R')

#std.pop.age.region.sex.race <- readRDS('./Data/std.pop.age.region.sex.race.rds')
std.pop.VA.age.region.sex.race <- readRDS('./Data/std.pop.VAenrollees.age.region.sex.race.rds')

```

Popsize for the CDC data
```{r}
#From Format_US_Population.R
#Format_US_Population()
filled_pop2 <- readRDS('./Data/pop_interpol.rds')

```

Setup CDC data and merge with popsize
```{r}
ag3 <- cdc_data_setup_multilevel()
ag3$log_pop <- log(ag3$pop.interpol/100000)

ag3 <- ag3[ag3$agec !="Under 25 Years" ,]
  # summary.cdc <- ag3[ag3$year<2020,] %>%
  #   group_by(agec, race_recode, region) %>%
  #   summarize(N_deaths=sum(N_deaths),pop=mean(pop.interpol, na.rm=T)) %>%
  #   mutate(inc_cdc=N_deaths/pop*100000) %>%
  #   select(agec, race_recode, region, inc_cdc)
  # 
  # saveRDS(summary.cdc,'./Data/grp_inc_cdc.R')

ag3$pandemic_effect <- 0
ag3$pandemic_effect[ag3$year>=2020 & ag3$qtr==1] <-1 
ag3$pandemic_effect[ag3$year>=2020 & ag3$qtr==2]  <-2
ag3$pandemic_effect[ag3$year>=2020 & ag3$qtr==3]  <-3
ag3$pandemic_effect[ag3$year>=2020 & ag3$qtr==4]   <- 4
ag3$pandemic_effect  <- as.factor(ag3$pandemic_effect )

```

Look at COVID deaths (based on ICD10) by group
```{r}
covid_counts <- readRDS('./Data/confidential/us_covid_deaths_by_group.rds') %>%
  mutate(month=(qtr*3 - 2), date=as.Date(paste(2020, month,'01', sep='-')))



  filled_pop2 <- readRDS('./Data/pop_interpol.rds') %>%
    mutate(sex = as.factor(sex), race_recode=as.factor(race_recode) , year=lubridate::year(date) ) %>%
    filter(year==2020) 

  



  covid_counts <- merge(covid_counts, filled_pop2, by.x=c('sex','race_recode','agec','region','date'),by.y=c('sex','race_recode','agec','region','date'), all.x=T)
  
  covid_age <- covid_counts %>%
    group_by(agec) %>%
    summarize(N_covid=sum(N_covid), pop=sum(pop.interpol)/4) %>%
    mutate(inc= N_covid/pop*100000)
  
 covid_total <- covid_counts %>%
   filter(!is.na(agec) & agec != 'Under 25 Years') %>%
    summarize(N_covid=sum(N_covid), pop=sum(pop.interpol)/4) %>%
    mutate(inc= N_covid/pop*100000)
  
```


https://www.cdc.gov/nchs/nvss/bridged_race/data_documentation.htm#vintage2020

Race:
Race:
RACE_RECODE: 
1=Non-Hispanic White
2=Non-Hispanic- Black
3= Hispanic
4=American Indian/Native Alaskan
5= Asian/Pacific Island
999=Missing


Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic

 01 ... Under 1 year (includes not stated infant ages)
 02 ... 1 - 4 years
 03 ... 5 - 14 years
 04 ... 15 - 24 years
 05 ... 25 - 34 years
 06 ... 35 - 44 years
 07 ... 45 - 54 years
 08 ... 55 - 64 years
 09 ... 65 - 74 years
 10 ... 75 - 84 years
 11 ... 85 years and over
 12 ... Age not stated




## Check inputs
Should be 1,012,805 deaths among 85+ year olds in 2020
1,013,340


In the US the death rate in 2019:
3,319,063 deaths among people age 25+;	226,635,013 people age 25+	1,464.5 deaths/100000
```{r}
N_deaths <- ag3[ag3$year==2020,] %>%
  group_by(agec) %>%
  summarize(N_deaths=sum(N_deaths))

```



## Fit Model


Fit to VA and CDC data separately. This function saves to an rds file summary tables with the excess incidence and RR by group (excluding raw counts and denominators). It also outputs the full summary tables that can be used in plotting in the list out.ds
```{r, eval=F}

out.ds <- call_inla(label='cdc', ds=ag3)

saveRDS(out.ds,'./outputs/out.ds.cdc.rds')
```

## Extract summary stats for table 2

```{r}
out.ds <-readRDS('./outputs/out.ds.cdc.rds')

results <- readRDS('./outputs/summary_list_cdc.rds')

extract_func(results, measure= 'RR', round.level = 2 )

extract_func(results, measure= 'RR_std', round.level = 2 )

extract_func(results, measure= 'excess_inc', round.level = 0 )

extract_func(results, measure= 'excess_std_inc', round.level = 0 )


extract_func(results, measure= 'pred_inc', round.level = 0 )

extract_func(results, measure= 'expected_std_inc', round.level = 0 )


extract_func(results, measure= 'expected_std_inc', round.level = 0 )


results$preds.overall$obs_inc
results$preds.overall$obs_std_inc

```




```{r, fig.width=8, fig.height=6}
out.ds$fit.plots$p1
out.ds$fit.plots$p3
out.ds$fit.plots$p5

```

```{r, fig.width=8, fig.height=6}
out.ds$rr.plot1$p2
out.ds$rr.plot1$p4
out.ds$rr.plot1$p6

```

```{r, fig.width=6, fig.height=2}

out.ds$preds.age.region.sex$agec <- as.character(out.ds$preds.age.region.sex$agec)

out.ds$preds.age.region.sex$agec <- factor(out.ds$preds.age.region.sex$agec, levels=c('25-44 years', '45-64 years','65-74 years', '75-84 years', '85 years and older'))

e1 <- out.ds$preds.age.region.sex[out.ds$preds.age.region.sex$region=='Northeast',]

ggplot(e1[e1$date>='2020-01-01',], aes(x=date, y=excess_inc_median, group=sex, color=sex)) +
  geom_line() +
  theme_classic() +
  facet_wrap(~agec , scales='free', nrow=1) +
  ylim(0,NA)

e2 <- out.ds$preds.age.region.sex[out.ds$preds.age.region.sex$region=='South',]

ggplot(e2[e2$date>='2020-01-01',], aes(x=date, y=excess_inc_median, group=sex, color=sex)) +
  geom_line() +
  theme_classic() +
  facet_wrap(~agec , scales='free', nrow=1)+
  ylim(0,NA)

e3 <- out.ds$preds.age.region.sex[out.ds$preds.age.region.sex$region=='West',]

ggplot(e3[e3$date>='2020-01-01',], aes(x=date, y=excess_inc_median, group=sex, color=sex)) +
  geom_line() +
  theme_classic() +
  facet_wrap(~agec , scales='free', nrow=1)+
  ylim(0,NA)

e4 <- out.ds$preds.age.region.sex[out.ds$preds.age.region.sex$region=='Midwest',]

ggplot(e4[e4$date>='2020-01-01',], aes(x=date, y=excess_inc_median, group=sex, color=sex)) +
  geom_line() +
  theme_classic() +
  facet_wrap(~agec , scales='free', nrow=1)+
  ylim(0,NA)
```

RR by race in the Northeast, by sex

```{r}
d1 <- out.ds$preds.age.race.region[out.ds$preds.age.race.region$region=='Northeast' ,]

 ggplot(d1, aes(x=date, y=RR_median, group=race_recode, col=race_recode)) +
   geom_line()+
   theme_classic()+
   facet_grid(sex~agec)


```

Sex differences by race and age in the northeast

```{r, fig.width=4, fig.height=4}
d1 <- out.ds$preds.age.race.region[out.ds$preds.age.race.region$region=='Northeast' ,]

d1$agec <- as.character(d1$agec)

d1$agec <- factor(d1$agec, levels=c('25-44 years', '45-64 years','65-74 years', '75-84 years', '85 years and older'))

 ggplot(d1, aes(x=date, y=RR_median, group=sex, col=sex)) +
   geom_line()+
   theme_classic()+
   facet_grid(race_recode~agec)

  ggplot(d1, aes(x=date, y=excess_inc_median, group=sex, col=sex)) +
   geom_line()+
   theme_classic()+
   facet_wrap(race_recode~agec, scales='free')

```

And in the South

```{r, fig.width=4, fig.height=4}
d1 <- out.ds$preds.age.race.region[out.ds$preds.age.race.region$region=='South' ,]

d1$agec <- as.character(d1$agec)

d1$agec <- factor(d1$agec, levels=c('25-44 years', '45-64 years','65-74 years', '75-84 years', '85 years and older'))

 ggplot(d1, aes(x=date, y=RR_median, group=sex, col=sex)) +
   geom_line()+
   theme_classic()+
       geom_hline(yintercept = 1)+

   facet_grid(race_recode~agec)

  ggplot(d1, aes(x=date, y=excess_inc_median, group=sex, col=sex)) +
   geom_line()+
   theme_classic()+
    geom_hline(yintercept = 0)+
   facet_wrap(race_recode~agec, scales='free')

```

## Compare excess and recorded covid
'non-COVID' means deaths not recorded as due to covid

RACE_RECODE: 
1=Non-Hispanic White
2=Non-Hispanic- Black
3= Hispanic
4=American Indian/Native Alaskan
5= Asian/Pacific Island
999=Missing

```{r, fig.width=6, fig.heigh=5}
detailed_summary <- results$preds.age.race.region %>%
  mutate(race_recode=as.numeric(as.character(race_recode))  ) %>%
  left_join(covid_counts, by=c('sex','race_recode', 'region','date', 'agec')) %>%
  mutate(      excess_deaths= excess_inc_median/100000*pop.interpol, 
               N_non_covid_excess=excess_deaths-N_covid,
               non_covid_excess_rate= N_non_covid_excess/pop.interpol*100000)

men <- detailed_summary %>%
  filter(sex=='M' & region=='South') %>%
  mutate(race_recode=as.factor(race_recode)) %>%
ggplot( aes(x=date, y=non_covid_excess_rate, group=race_recode, col=race_recode, lty=race_recode)) +
  geom_line() +
  theme_classic() +
  ylab('Excess non-COVID deaths/100000')+
  facet_wrap(~agec, scales='free')
men


aggr_summary <- detailed_summary %>%
  group_by(agec,race_recode, date) %>%
    mutate(race_recode=factor(race_recode, c('1','2','3','4','5'), c('White','Black' ,'Hispanic','Amer.Indian','Asian/Paci/Is'))) %>%

  summarize(N_non_covid_excess=sum(N_non_covid_excess), pop.interpol=sum(pop.interpol)) %>%
    mutate( non_covid_excess_rate= N_non_covid_excess/pop.interpol*100000) %>%
ggplot( aes(x=date, y=non_covid_excess_rate, group=race_recode, col=race_recode, lty=race_recode)) +
  geom_line() +
  theme_classic() +
  ylab('Excess non-COVID deaths/100000')+
  facet_wrap(~agec, scales='free')
aggr_summary
```



