---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "10/29/2020"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
params:
  agg.level: 'state'
  n.days.filter: 20
  web.version: FALSE
  extrap.date: '2020-01-26'
  count.start.date: '2020-03-01'
  end.data.date: '2020-09-05'
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

```

```{r setup}
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(parallel)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(patchwork)
library(abind)
library(INLA)
library(gsubfn)
library(dplyr)
library(RCurl)
library(dplyr)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(rjags)
library(INLA)
library(lme4)
library(pbapply)
#library(jsonlite)
set.seed(123)

source('./R/process_va_region.R')
source('./R/process_nchs_region.R')
source('./R/inla_mod_grp.R')
source('./R/plot_ts_region.R')
source('./R/run_if_expired.R')
source('./R/gen_pred_interval_inla.R')
source('./R/popsize_formatting.R')
source('./R/setup_CDC data_multilevel.R')
source('./R/inla_model1.R')
```

Popsize for the CDC data
```{r}
filled_pop2 <- census_popsize_format()

```

Setup CDC data and merge with popsize
```{r}
ag3 <- cdc_data_setup_multilevel()
ag3$log_pop <- log(ag3$pop.interpol/100000)
```

Read in VA data

```{r}
v1 <- read_csv('./Data/Confidential/aggdata_withregion.csv')
v1$race_recode[is.na(v1$race_recode)] <- 999
v1$pop <- v1$n #pop is specific to the agec, region, sex, year but does NOT vary by race or quarter

v1$N_deaths[v1$N_deaths=='<10'] <- 5 #just as a placeholder...can either have liam run the analysis locally to avoid masking, or figure out an imputation method
v1$N_deaths <- as.numeric(v1$N_deaths)

  v1$month <- NA
  v1$month[v1$qtr==1] <- 1
  v1$month[v1$qtr==2] <- 4
  v1$month[v1$qtr==3] <- 7
  v1$month[v1$qtr==4] <- 10
  v1$date <- as.Date(paste(v1$year, v1$month, '01', sep='-'))

  v1 <- v1[v1$date<=as.Date('2021-01-01'),]
  
  v1 <- v1[!is.na(v1$region) & !is.na(v1$sex) & !is.na(v1$agec),] # for now, ignore deaths where sex, or region or age is NA
  
  v1 %>%
    group_by(agec, race_recode, region) %>%
    summarize(N_deaths=sum(N_deaths),pop=mean(n))


  v1$source <- 'va'
  
  v1$time <- round(as.numeric(v1$date - min(v1$date))/30.3)
  v1$time_scale <- as.vector(scale(v1$time)) 

  
   v1$qtr2 <- 0
  v1$qtr2[v1$qtr==2] <- 1
  v1$qtr3 <- 0
  v1$qtr3[v1$qtr==3] <- 1
  v1$qtr4 <- 0
  v1$qtr4[v1$qtr==4] <- 1
  
    v1$subgroup_combo <- as.factor(paste(v1$race_recode, v1$agec, v1$sex,v1$region))

      v1$all_cause_pre <- v1$N_deaths
  v1$all_cause_pre[v1$date >= '2020-01-01'] <- NA #bc it is quarterly data, Q1 includes pandeic period
  v1$race_recode <- as.factor(v1$race_recode)
  
  v1 <- v1[v1$agec !="Under 25 Years",]

    v1$inc <- v1$N_deaths/v1$pop*100000
    v1$log_pop <- log(v1$pop/100000)
    
    v1 <- v1[v1$year <=2020,]
```


https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-detail.html
Race:
1 = White Alone
2 = Black or African American Alone
3 = American Indian or Alaska Native Alone
4 = Asian Alone
5 = Native Hawaiian and Other Pacific Islander Alone
6 = Two or more races

Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic

 01 ... Under 1 year (includes not stated infant ages)
 02 ... 1 - 4 years
 03 ... 5 - 14 years
 04 ... 15 - 24 years
 05 ... 25 - 34 years
 06 ... 35 - 44 years
 07 ... 45 - 54 years
 08 ... 55 - 64 years
 09 ... 65 - 74 years
 10 ... 75 - 84 years
 11 ... 85 years and over
 12 ... Age not stated



## Model

```{r}


#Try different levels of tau to find optimum...this is a compromise between doing a full 
test.tau <-  c(0.01,0.1,0.4,0.5,0.6,0.7, 0.7,1,1.1, 1.25, 1.5,2)
mod.comp.va <- pblapply( test.tau, function(x) inla_model1(ds=v1, prec.prior1=x))
mod.comp.cdc <- pblapply( test.tau, function(x) inla_model1(ds=ag3, prec.prior1=x))

waic.va <- cbind.data.frame('test.tau'=test.tau, 'waic'=sapply(mod.comp.va, '[[','waic'))
dic.va <- cbind.data.frame('test.tau'=test.tau, 'dic'=sapply(mod.comp.va, '[[','dic'))

plot(dic.va$test.tau, dic.va$dic-min(dic.va$dic), type='l')
abline(h=2)

```



Fit to VA and CDC data separately
```{r}

combine.data <- list('cdc'=ag3, 'va'=v1)

mod1 <- pblapply(combine.data, inla_model1)


res1 <- bind_rows(mod1$cdc$preds_covars.m, mod1$va$preds_covars.m) #add VA data too

```

We can then take the results from the INLA and aggregate over different groups. For example, by age:
```{r}


preds.age <- res1 %>%
  group_by(source,agec, date, variable) %>%
  summarise( 'N_deaths'=sum(N_deaths), 'pred'=sum(value) ) %>%
  ungroup() %>%
  group_by(source,agec, date) %>%
  summarise( 'N_deaths'=mean(N_deaths), 'pred_median'=median(pred), pred_lcl=quantile(pred,probs=0.025), pred_ucl=quantile(pred,probs=0.975) )
```

```{r, fig.width=8, fig.height=6}
 p1 <- ggplot(preds.age[preds.age$source=='va',], aes(x=date, y=N_deaths , group=agec, col=agec)) +
    geom_line() +
    geom_line(aes(x=date, y=pred_median, group=agec), lty=2, col='black') +
    geom_ribbon(aes(ymin=pred_lcl, ymax=pred_ucl), alpha=0.2, col='gray') +
    ylab("Number of all-cause deaths") +
    xlab("Date") +
    theme_classic() +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
    geom_hline(yintercept=0, col='gray', lty=2) +
    geom_vline(xintercept=as.Date('2020-03-01'), col='gray', lty=2) +
        facet_wrap(~agec , scales='free') +
    xlim(as.Date('2014-01-01'), as.Date('2020-12-31')) +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 
p1
```

or by age and race
```{r}
preds.age.race <- res1 %>%
  group_by(source,agec,race_recode, date, variable) %>%
  summarise( 'N_deaths'=sum(N_deaths), 'pred'=sum(value) ) %>%
  ungroup() %>%
  group_by(source,agec, race_recode,date) %>%
  summarise( 'N_deaths'=mean(N_deaths), 'pred_median'=median(pred), pred_lcl=quantile(pred,probs=0.025), pred_ucl=quantile(pred,probs=0.975) )
```

```{r}
 p1 <- ggplot(preds.age.race[preds.age.race$agec=='65-74 years' & preds.age.race$source=='cdc',], aes(x=date, y=N_deaths )) +
    geom_line() +
    geom_line(aes(x=date, y=pred_median, group=agec), lty=2, col='black') +
    geom_ribbon(aes(ymin=pred_lcl, ymax=pred_ucl), alpha=0.2, col='gray') +
    ylab("Number of all-cause deaths") +
    xlab("Date") +
    theme_classic() +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
    geom_hline(yintercept=0, col='gray', lty=2) +
    geom_vline(xintercept=as.Date('2020-03-01'), col='gray', lty=2) +
        facet_wrap(~race_recode , scales='free') +
    xlim(as.Date('2014-01-01'), as.Date('2020-12-31')) +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 
p1
```