---
title: "Excess death in the US by age"
author: "Dan Weinberger"
date: "10/29/2020"
output:
  html_document:
    df_print: paged
    html_document: null
    toc: yes
    toc_depth: 2
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document: 
    keep_tex:  true
params:
  agg.level: 'state'
  n.days.filter: 20
  web.version: FALSE
  extrap.date: '2020-01-26'
  count.start.date: '2020-03-01'
  end.data.date: '2020-09-05'
---

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo=F,
  warning=FALSE, 
  message=FALSE,
  comment = "#>",
  dev=c('png','pdf'),
  fig.path='./figures/',
  gganimate = list(
    nframes = 50)
)

extrap.date <-  as.Date(params$extrap.date)
count.start.date <- as.Date(params$count.start.date)
end.data.date <- as.Date(params$end.data.date)

state.name2 <- c(state.name, 'District of Columbia','Puerto Rico', 'United States', 'New York City')
state.abb2 <- c(state.abb, 'DC','PR','US','NYC')


last.date.format <- 
  format(end.data.date, '%b %d, %Y')
```

```{r setup}
library(ExcessILI)
library(readxl)
library(cdcfluview)
library(reshape2)
library(ggplot2)
library(lubridate)
library(RColorBrewer)
library(plotly)
library(MMWRweek)
library(readr)
library(rjson)
library(htmlTable)
library(RSocrata)
library(pdftools)
library(readr)
library(abind)
library(gsubfn)
library(dplyr)
library(RCurl)
library(tidyverse)
library(gifski)
library(gganimate)
library(shiny)
library(rjags)
library(HDInterval)
library(INLA)
library(stringr)
library(lme4)
#library(jsonlite)
set.seed(123)
source('./R/gen_pred_interval_inla.R')

```

## Population size data
https://www.census.gov/programs-surveys/popest/technical-documentation/research/evaluation-estimates/2020-evaluation-estimates/2010s-state-detail.html
Race:
1 = White Alone
2 = Black or African American Alone
3 = American Indian or Alaska Native Alone
4 = Asian Alone
5 = Native Hawaiian and Other Pacific Islander Alone
6 = Two or more races

Sex: 
0 = Total
1 = Male
2 = Female

The key for ORIGIN is as follows:
0 = Total
1 = Not Hispanic
2 = Hispanic

 01 ... Under 1 year (includes not stated infant ages)
 02 ... 1 - 4 years
 03 ... 5 - 14 years
 04 ... 15 - 24 years
 05 ... 25 - 34 years
 06 ... 35 - 44 years
 07 ... 45 - 54 years
 08 ... 55 - 64 years
 09 ... 65 - 74 years
 10 ... 75 - 84 years
 11 ... 85 years and over
 12 ... Age not stated

## Set up data for multilevel model

```{r}

ag3 <- readRDS('./Data/Confidential/compiled_sex_age_race_qtr.rds')
ag3$month <- NA
ag3$month[ag3$qtr==1] <- 1
ag3$month[ag3$qtr==2] <- 4
ag3$month[ag3$qtr==3] <- 7
ag3$month[ag3$qtr==4] <- 10
ag3$date <- as.Date(paste(ag3$year, ag3$month, '01', sep='-'))

ag3$qtr2 <- 0
ag3$qtr2[ag3$qtr==2] <- 1

ag3$qtr3 <- 0
ag3$qtr3[ag3$qtr==3] <- 1

ag3$qtr4 <- 0
ag3$qtr4[ag3$qtr==4] <- 1


#ag3 <- ag3[ag3$state!='US',]

ag3$time <- round(as.numeric(ag3$date - min(ag3$date))/30.3)

ag3$age_group <- as.factor(ag3$agec)
ag3$race_recode <- as.factor(ag3$race_recode)
ag3$sex <- as.factor(ag3$sex)

#ag3$state <- as.factor(ag3$state)

#mod.inla <- inla(all_cause ~  , family='poisson')

ag3$all_cause_pre <- ag3$N_deaths

ag3$all_cause_pre[ag3$date >= '2020-03-01'] <- NA
ag3$time_scale <- as.vector(scale(ag3$time))

# ag3.pre <- ag3[ag3$week_end < '2020-03-01',] %>%
#   group_by(state,age_group) %>%
#   summarize( log.ave.case = log(mean(all_cause)  ))
  
#ag3 <- merge( ag3, ag3.pre, by=c('state','age_group'))



ag3 <- merge(ag3, filled_pop2, by.x=c('sex','race_recode','agec', 'date'),by.y=c('sex','race','agec', 'date'), all.x=T)


ag3$log_pop <- log(ag3$pop.interpol)

ag3$subgroup_combo <- as.factor(paste(ag3$race_recode, ag3$agec, ag3$sex))
ag3$subgroup_combo2 <- ag3$subgroup_combo
ag3$subgroup_combo3 <- ag3$subgroup_combo
ag3$subgroup_combo4 <- ag3$subgroup_combo
ag3$subgroup_combo5 <- ag3$subgroup_combo

ag3$agec2 <- ag3$agec
ag3$agec3 <- ag3$agec
ag3$agec4 <- ag3$agec
ag3$agec5 <- ag3$agec

ag3$race_recode2 <- ag3$race_recode
ag3$race_recode3 <- ag3$race_recode
ag3$race_recode4 <- ag3$race_recode
ag3$race_recode5 <- ag3$race_recode

ag3$sex2 <- ag3$sex
ag3$sex3 <- ag3$sex

#ag3$state2 <- ag3$state
#ag3$state2 <- ag3$state

ag3 <- ag3[!is.na(ag3$pop.interpol),]

ag3$obs <- as.factor(1:nrow(ag3))

#ag3 <- ag3[ag3$year<=2020,]
```

## Multilevel model

Simple GLM
```{r}
mod.glm1 <- glm(all_cause_pre ~ 
                1 + time_scale + qtr2 + qtr3 +qtr4 +
                  agec + race_recode + sex + subgroup_combo +
                  agec*(time_scale + qtr2 + qtr3 +qtr4) +
                  race_recode*(time_scale + qtr2 + qtr3 +qtr4)  
                  #sex*(time_scale + sin1 + sin2 + sin3 + cos1 +cos2+ cos3) + 
                     ,
                offset = log_pop,
                family='poisson', data=ag3)
summary(mod.glm1)
```
```{r}
library(rstanarm)

mod.lme42 <- stan_glmer(all_cause_pre ~ 
                1 + time_scale +qtr2 + qtr3 +qtr4 +
                   sex  +(1+time_scale +qtr2 + qtr3 +qtr4|agec) + (1+time_scale +qtr2 + qtr3 +qtr4|race_recode), 
                offset = log_pop,
                data=ag3,
                family='poisson')

```


Model fit takes about 1 min
```{r}

mod.inla2 <- inla(all_cause_pre ~
              1 + time_scale + qtr2 + qtr3 +qtr4 +
                  agec + race_recode + sex + subgroup_combo +
                  agec*(time_scale + qtr2 + qtr3 +qtr4) +
                  race_recode*(time_scale + qtr2 + qtr3 +qtr4)  +
                  sex*(time_scale + qtr2 + qtr3 +qtr4)                                       ,
                                offset = log_pop,
                family='nbinomial', data=ag3,
                control.compute=list(config = TRUE))
summary(mod.inla2)


preds.inla2 <- gen_pred_interval_inla(mod.inla2, ag3)


# prec.prior <- list(prec = list(param = c(0.001, 0.001)))
# 
# test.mod <- inla(all_cause_pre ~
#               1 + time_scale + qtr2 + qtr3 +qtr4 +
#                     sex  +
#                 f(agec, model='iid',constr = FALSE,hyper = prec.prior) +
#                 f(agec2, time_scale, model='iid', hyper = prec.prior) +
#                  f(agec3, qtr2, model='iid',hyper = prec.prior) +
#                  f(agec4, qtr3, model='iid', hyper = prec.prior) +
#                  f(agec5, qtr4, model='iid', hyper = prec.prior) 
# 
#                 # f(race_recode, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(race_recode2, time_scale, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(race_recode3, qtr2, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(race_recode4, qtr3, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(race_recode5, qtr4, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # 
#                 # f(subgroup_combo, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(subgroup_combo2, time_scale, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(subgroup_combo3, qtr2, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(subgroup_combo4, qtr3, model='iid',scale.model = TRUE, hyper = prec.prior) +
#                 # f(subgroup_combo5, qtr4, model='iid',scale.model = TRUE, hyper = prec.prior)
# 
#               ,
#                                 offset = log_pop,
#                 family='nbinomial', data=ag3,
#                 control.compute=list(config = TRUE))
# preds.inla2 <- gen_pred_interval_inla(test.mod, ag3)



```



We can then take the results from the INLA and aggregate over different groups. For example, by age:
```{r}
preds.age <- preds.inla2$preds_covars.m %>%
  group_by(agec, date, variable) %>%
  summarise( 'N_deaths'=sum(N_deaths), 'pred'=sum(value) ) %>%
  ungroup() %>%
  group_by(agec, date) %>%
  summarise( 'N_deaths'=mean(N_deaths), 'pred_median'=median(pred), pred_lcl=quantile(pred,probs=0.025), pred_ucl=quantile(pred,probs=0.975) )


```

```{r, fig.width=8, fig.height=12}


 p1 <- ggplot(preds.age, aes(x=date, y=N_deaths , group=agec, col=agec)) +
    geom_line() +
    geom_line(aes(x=date, y=pred_median, group=agec), lty=2, col='black') +
    geom_ribbon(aes(ymin=pred_lcl, ymax=pred_ucl), alpha=0.2, col='gray') +
    ylab("Number of all-cause deaths") +
    xlab("Date") +
    theme_classic() +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
    geom_hline(yintercept=0, col='gray', lty=2) +
    geom_vline(xintercept=as.Date('2020-03-01'), col='gray', lty=2) +

        facet_wrap(~agec , scales='free') +
    xlim(as.Date('2014-01-01'), as.Date('2020-12-31')) +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 

p1
```
JAGS model
```{r}
time.vars <- unique(cbind.data.frame('date'=ag3$date, 'qtr2'=ag3$qtr2, 'qtr3'=ag3$qtr3, 'qtr4'=ag3$qtr4, 'time_Scale'=ag3$time_scale))

ag3.m <- reshape2::melt(ag3[,c('N_deaths','log_pop','agec', 'date','race_recode','sex')], id.vars=c('agec','race_recode','sex','date') )

ag3.c <- acast( ag3.m, variable~agec~race_recode~sex~date)

source('./R/jags_mod1_non_nested.R')

ag4 <- ag3[,c('N_deaths','log_pop','agec', 'date','race_recode','sex')]
ag4$race.index <- as.numeric(ag4$race_recode)
ag4$agec.index <- as.numeric(as.factor(ag4$agec))
ag4$sex.index <- as.numeric(as.factor(ag4$sex))
ag4$time.index <- as.numeric(as.factor(ag4$date))

#table(ag4$date, ag4$time.index) #check assignments

post_preds2 <- merge(post_preds, ag4, by.x=c('agec','race','sex','time'), by.y=c('agec.index','race.index','sex.index','time.index'))


sample_key <- 
pos_samples <- cbind.data.frame(coords.df)
```

```{r}
p1 <- ggplot(post_sample_age, aes(x=time, y=pred_median , group=agec, col=agec)) +
    geom_line() +
    geom_line(aes(x=time, y=pred_median, group=agec), lty=2, col='black') +
    geom_ribbon(aes(ymin=pred_lcl, ymax=pred_ucl), alpha=0.2, col='gray') +
    ylab("Number of all-cause deaths") +
    xlab("Date") +
    theme_classic() +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) +
    geom_hline(yintercept=0, col='gray', lty=2) +
   # geom_vline(xintercept=as.Date('2020-03-01'), col='gray', lty=2) +

        facet_wrap(~agec , scales='free') +
   # xlim(as.Date('2014-01-01'), as.Date('2020-12-31')) +
    theme(panel.spacing= unit(2,'lines') , axis.text.x=element_text(angle=90)) 

p1
```

